---
title: "Math 327 Chapter 4 Homework"
output: word_document
---

### __ Your name here __

4.3a  __ Type answer here __

```{r}
# Code for 4.3b
# Open the data file, CH01PR2.txt
mydata <- read.table(file="__ add your full path here__/CH01PR20.txt",header=F,col.names=c("Y","X"))

xname = "Service Time (minutes)"
yname = "Copiers Serviced (#)"

attach(mydata)

myfit <- lm (Y ~ X)
myfit
confint(myfit, level= __set level__)   

```

4.3b __ State the joint confidence limits with b0 and b1 here. __

4.3c. __ Type answer here __

```{r}
# Fit a regression through the origin, for 4.16a
int0fit = lm (Y ~ 0 + X)
int0fit
```

4.16a. __ State the regression function here __

```{r}
confint(int0fit, level=0.90)
```

4.16b.  __ State your answer here __

```{r}
predict(int0fit, data.frame(X=c(6)), interval="prediction", level= __set level__)
```

4.16c.  __ State your answer here __

```{r}
# Plot the data
plot.new()
plot(X, Y, xlab=xname, ylab=yname, main="Regression Through the Origin")
abline(myfit)
```

4.17a. __ State your answer here __


```{r}
# save the residuals
int0resid = int0fit$residuals
sum (int0resid)

# Plot residuals vs fitted
plot.new()
plot (int0fit$fitted.values, int0resid, xlab=paste(yname, "fitted values"), ylab="Residuals", main="Regression Through Origin - Residuals vs Fitted")
abline(h=0)
```

4.17b.  __ State your answer here __

```{r}
# Lack of fit test
full = lm (Y ~ 0 + as.factor(X))
anova(int0fit, full)
```

4.17c.  Lack of fit test.  State hypotheses, decision rule (can be in terms of p-value), and conclusion. 


```{r}
# inverse prediction using the GPA data.
# Need to read in the GPA data, and change variable names.
mydata2 <- read.table(file="__insert your path here__/CH01PR19.txt",header=F,col.names=c("Y","X"))
xname = "ACT"
yname = "GPA"
attach(mydata2)
myfit = lm (Y ~ X)
plot (X, Y, xlab=xname, ylab=yname, main="Inverse Prediction")
abline (myfit)
b0 = myfit$coeff[1]
b1 = myfit$coeff[2]
xnew = (3.4 - b0)/b1
summ = summary (myfit)
mse = summ$sigma^2
n = length(Y)
numer = (xnew - mean(X))^2
denom = sum ((X - mean(X))^2)
s.predx.sq = (mse/b1^2)*(1 + 1/n + numer/denom)
xnew.lower = xnew - qt(0.90, n-2)*sqrt(s.predx.sq)  # Change confidence level, as needed
xnew.upper = xnew + qt(0.90, n-2)*sqrt(s.predx.sq)  # Change confidence level, as needed

# Prediction and Confidence interval for predicted X value at Y=16
data.frame(Xnew = c(xnew), Lower = c(xnew.lower), Upper = c(xnew.upper), row.names=c("Prediction"))

```

4.19a __ State your answer here. include interpretaiton.  __

```{r}
# Equation to check the reasonableness of the interval, equation 4.33, p. 170 should be approx. < 0.1
eqn4.33 = (qt(0.90, n-2)^2)*mse/(b1^2 * denom)
eqn4.33
```

4.19b __ State your answer here __

